Poincar\'e--Friedrichs constants over triangulated domains

### 2024-06-15 draft.tex
add makefile; extensions and polishment
work on combinatorics

>>>>

%  
 % \color{red}We want to summarize the Lp norms so that the Poincare inequality over $U_F$ is applied only once. In the special cases $p=1$ and $p=\infty$, this works as above. Do we have a formula for general $p$ that contains these as limit cases? \color{black}
 % In the special case $p < \infty$, we use the finite-dimensional H\"older inequality to find 
 % \begin{align*}
 %    \| w_{m} \|_{L^{p}(T_m)}^{p}
 %    &\leq 
 %    3^{p-1}
 %    \| u_{F_m} \|_{L^{p}(T_m)}^{p}
 %    +
 %    3^{p-1}
 %    \frac{ \vol(T_m) }{ \vol(T_{j(m)}) }
 %    \| u_{F_{m}} \|_{L^{p}(T_{j(m)})}^{p}
 %    +
 %    3^{p-1}
 %    \frac{ \vol(T_m) }{ \vol(T_{j(m)}) }
 %    \| w_{m-1}\|_{L^{p}(T_{j(m)})}^{p}
 %    \\&\leq 
 %    3^{p-1}
 %    \max\left( 1, \frac{ \vol(T_m) }{ \vol(T_{j(m)}) } \right)
 %    \| u_{F_m} \|_{L^{p}(U_{F_m})}^{p}
 %    +
 %    3^{p-1}
 %    \frac{ \vol(T_m) }{ \vol(T_{j(m)}) }
 %    \| w_{m-1}\|_{L^{p}(T_{j(m)})}^{p}
 %    ,
 % \end{align*}
 % and therefore
 % \begin{align*}
 %    \| w_{m} \|_{L^{p}(T_m)}
 %    &\leq 
 %    3^{\frac{p-1}{p}}
 %    \max\left( 1, \frac{ \vol(T_m) }{ \vol(T_{j(m)}) } \right)^{\frac 1 p}
 %    \| u_{F_m} \|_{L^{p}(U_{F_m})} 
 %    +
 %    3^{\frac{p-1}{p}}
 %    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{j(m)})^{\frac 1 p} }
 %    \| w_{m-1}\|_{L^{p}(T_{j(m)})} 
 %    .
 % \end{align*}
 % \color{blue}
 % In the special case $p < \infty$, we use the finite-dimensional H\"older inequality to find 
 % \begin{align*}
 %    \| w_{m} \|_{L^{p}(T_m)}^{p}
 %    &\leq 
 %    \left( 
 %        1 + 2 \frac{ \vol(T_m)^{\frac{1}{p-1}} }{ \vol(T_{j(m)})^{\frac{1}{p-1}} }
 %    \right)^{ p-1 }
 %    \left( 
 %        \| u_{F_m} \|_{L^{p}(T_m)}^{p}
 %        +
 %        \| u_{F_{m}} \|_{L^{p}(T_{j(m)})}^{p}
 %        +
 %        \| w_{m-1}\|_{L^{p}(T_{j(m)})}^{p}
 %    \right)
 %    ,
 % \end{align*}
 % and therefore
 % \begin{align*}
 %    \| w_{m} \|_{L^{p}(T_m)}
 %    &\leq 
 %    \left( 
 %        1 + 2 \frac{ \vol(T_m)^{\frac{1}{p-1}} }{ \vol(T_{j(m)})^{\frac{1}{p-1}} }
 %    \right)^{ \frac{p-1}{p} }
 %    \left( 
 %        \| u_{F_m} \|_{L^{p}(T_m)} 
 %        +
 %        \| u_{F_{m}} \|_{L^{p}(T_{j(m)})} 
 %        +
 %        \| w_{m-1}\|_{L^{p}(T_{j(m)})} 
 %    \right)
 %    .
 % \end{align*}
 % \color{black}

we have any of the following two conditions:
\begin{itemize}
 \item 
 there exists $\epsilon > 0$ 
 and a homeomorphism 
 \begin{align*}
    \phi : \Ball_\epsilon(x) \cap \underlying{\calT}  
    \rightarrow 
    \left\{ x \in \bbR^{n} \suchthat |x| \leq 1 \right\}
 \end{align*}
 such that $\phi(x) = 0$,
 \item 
 there exists $\epsilon > 0$ and a homeomorphism 
 \begin{align*}
    \phi : \Ball_\epsilon(x) \cap \underlying{\calT} 
    \rightarrow 
    \left\{ x \in \bbR^{n} \suchthat |x| \leq 1, x_n \leq 0 \right\}
 \end{align*}
 such that $\phi(x) = 0$.
\end{itemize}
This characterizes the union of the simplicial complex $\calT$ as a topological manifold with boundary.

\subsection{On the topology of shellable triangulations}

A shellable simplicial complex that triangulates a manifold has a very restricted topology. 
The Mayer-Vietoris theorem implies restrictions on the homology groups of simplicial complexes iteratively constructed via shelling.

    Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, 
    whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. 
    In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. 
    We write $\calT'$ for the  simplicial complex that contains $\calT \cup T$.

Suppose that $\calT$ is a strongly connected $n$-dimensional simplicial complex that triangulates a manifold, and suppose that its boundary complex $\partial\calT$ is either empty or strongly connected.  
Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. We write $\calT'$ be the  simplicial complex that contains $\calT \cup T$.
We exclude the case $n \leq 1$ since otherwise the discussion is trivial.

According to the simplicial Mayer-Vietoris theorem, there exists an exact sequence 
\begin{align*}
    \begin{CD}
        \dots \longrightarrow H_{k}(\calS) \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow H_{k-1}(\calS) \longrightarrow\dots 
    \end{CD}
\end{align*}
We notice that $\calS$, $\calT$, $T$ and $\calT'$ are each connected, and hence their zeroth homology groups are isomorphic to ${G}$. 
By the nature of the mappings in the Mayer-Vietoris sequence, 
\color{red}which uses some specific knowledge about the arrows\color{black}, 
we see that 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
is an exact sequence. 
Moreover, $H_{k}(\calS) = H_{k}(T) = 0$ for $0 < k < n-1$. Hence, for $1 < k < n-1$,
the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
and we conclude that $H_{k}(\calT) = 0$ for $0 < k < n-1$.
Finally, the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{n}(\calT \oplus T) \longrightarrow H_{n}(\calT') \longrightarrow H_{n-1}(\calS) \longrightarrow H_{n-1}(\calT \oplus T) \longrightarrow H_{n-1}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}

% Here, if $n=1$, it already follows that $H_{n}(\calT \oplus T) \simeq H_{n}(\calT')$.
% % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq \bbZ$ because the respective simplicial complexes are path-connected. 
% Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
% one easily sees that $H_{0}(\calT') \simeq \bbZ$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
Since $\calT$ and $T$ are triangulations of manifolds with non-empty boundary, their $n$-th simplicial homology groups vanish. 
then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ are trivial and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq \bbZ$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

In the case that $\calS$ is isomorphic to a disk, we have $H_{n-1}(\calS) = 0$,
and then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T) = 0$ are trivial 
and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq {G}$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

We conclude that the simplicial homology groups of a shellable $n$-dimensional simplicial complex $\calT$ that triangulates a manifold satisfy 
\begin{align*}
    H_0(\calT) \simeq \bbZ, 
    \quad 
    H_1(\calT) \simeq \dots \simeq H_{n-1}(\calT) \simeq 0.
\end{align*}
Moreover, $H_n(\calT) \simeq {G}$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise. 
Moreover, $H_n(\calT) \simeq \bbZ$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise.

% then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follow immediately.
% Moreover, Thus $H_{n}(\calT') = H_{n}(\calT) \oplus H_{n}(\calS)$.
% We conclude that any shellable simplicial complex has the homology groups of a disk or a sphere. 
% NB: the zeroth homology is always free, and the n-th homology group, being the kernel of a mapping free abelian group, is free as well.
% \begin{align*}
%     \begin{CD}
%         0 \longrightarrow H_{1}(\calT) \oplus H_{1}(T) \longrightarrow H_{1}(\calT') \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
%     \end{CD}
% \end{align*}

\begin{remark}    
    % % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq {G}$ because the respective simplicial complexes are path-connected. 
    % Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
    % one easily sees that $H_{0}(\calT') \simeq {G}$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$.

    % then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follow immediately.
    % Moreover, Thus $H_{n}(\calT') = H_{n}(\calT) \oplus H_{n}(\calS)$.
    % We conclude that any shellable simplicial complex has the homology groups of a disk or a sphere. 
    % NB: the zeroth homology is always free, and the n-th homology group, being the kernel of a mapping free abelian group, is free as well.
    % \begin{align*}
    %     \begin{CD}
    %         0 \longrightarrow H_{1}(\calT) \oplus H_{1}(T) \longrightarrow H_{1}(\calT') \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    %     \end{CD}
    % \end{align*}
\end{remark}

### 2024-06-15 main.tex
work on combinatorics

Given any simplex $T$, we write $\subsimplex(T)$ for the simplicial complex that contains all subsimplices of $T$. 
More specifically, $\subsimplex_{{k}}(T) \subseteq \subsimplex(T)$ denotes the set of $k$-dimensional subsimplices of $T$. 
Whenever $\calT$ is a simplicial complex and $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$.
We write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$. 
Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. 
In practice, we not always distinguish between points and singleton simplices. % TODO: see whether being formally correct iis possible 
Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (codimension $1$ members) of this triangulation.

@@

@@

\begin{lemma}    
    Any simplicial $2$-disk is shellable.
    Any simplicial $2$-sphere is shellable. 
\end{lemma}
\begin{proof}
    Let $\calT$ be a simplicial $2$-disk. 
    If $\calT$ contains only one triangle, then there is nothing to show. 
    Otherwise, there exists at least one triangle $T$ that touches the boundary of the disk at either one or two edges 
    {\color{red}Needs proof}. 
    Letting $\calT'$ be the triangulation obtained by removing $T$,
    we have another triangulation of the $2$-disk that intersects $T$ only at either two or one edges.
    Repeating this procedure yields an enumeration of the triangles of $\calT$,
    whose reverse is a shelling.

    Now let $\calS$ be any simplicial $2$-sphere. 
    By removing any triangle $S \in \calS$, we obtain a simplicial $2$-disk, which is shellable. 
    Any of its shellings can be extended to a shelling of $\calS$ by re-inserting the first triangle $S$.
\end{proof}

@@

    It remains to consider the case where $S$ is a boundary vertex. 
    Then the triangles of $\patch_{\calT}(S)$ that do not contain $V$ constitute a simplicial $2$-disk 
    {\color{red}Needs proof},
    which is shellable,
    and any such shelling can be extended to a shelling of $\patch_{\calT}(S)$.
\end{proof}

\begin{example}
    Not all trianguable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler2012lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions. 
\end{example}
\begin{remark}
    Patches around boundary simplices are not necessarily shellable when the dimension exceeds $3$.
    For example, there exists a non-shellable triangulation of a tetrahedron. 
    If we extend this triangulation to a triangulation of a hypertetrahedron by suspending it from a new point $v_\star$, 
    then the resulting new triangulation is non-shellable and coincides with the patch around $v_\star$.
\end{remark}

The reason of our interest in shellable simplicial complexes is that they can be constructed by successive adhesion of simplices to intermediate complexes, 
such that the interface is well-behaved: it is a simplicial disk or a simplicial sphere. 
That has the following important consequence.

% TODO: we only need some basic properties, namely that the construction maintains manifolds with boundary.
% TODO: copy and adapt the content of Ziegler and Kozlov as necessary

\begin{lemma}
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots$, then it is a manifold triangulation.
    In particular, $\patch_{\calT}(T_{0}) \cup \patch_{\calT}(T_{1}) \cup \dots \cup \patch_{\calT}(T_{m})$ is a manifold triangulation for every $m \geq 0$.
\end{lemma}
\begin{proof}    
\end{proof}

\begin{lemma}    
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots$ and is non-branching, then it triangulates a manifold, possibly with boundary.
    In particular, $T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a manifold with boundary for all $m \geq 0$.
\end{lemma}
\begin{proof}    
\end{proof}

% The notion of shelling is taken from Ziegler's book~\cite{ziegler2012lectures}

\subsection{On the topology of shellable triangulations}

A shellable simplicial complex that triangulates a manifold has a very restricted topology. 
The Mayer-Vietoris theorem implies restrictions on the homology groups of simplicial complexes iteratively constructed via shelling.

    Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, 
    whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. 
    In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. 
    We write $\calT'$ for the  simplicial complex that contains $\calT \cup T$.

Suppose that $\calT$ is a strongly connected $n$-dimensional simplicial complex that triangulates a manifold, and suppose that its boundary complex $\partial\calT$ is either empty or strongly connected.  
Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. We write $\calT'$ be the  simplicial complex that contains $\calT \cup T$.
We exclude the case $n \leq 1$ since otherwise the discussion is trivial.

According to the simplicial Mayer-Vietoris theorem, there exists an exact sequence 
\begin{align*}
    \begin{CD}
        \dots \longrightarrow H_{k}(\calS) \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow H_{k-1}(\calS) \longrightarrow\dots 
    \end{CD}
\end{align*}
We notice that $\calS$, $\calT$, $T$ and $\calT'$ are each connected, and hence their zeroth homology groups are isomorphic to ${G}$. 
By the nature of the mappings in the Mayer-Vietoris sequence, 
\color{red}which uses some specific knowledge about the arrows\color{black}, 
we see that 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
is an exact sequence. 
Moreover, $H_{k}(\calS) = H_{k}(T) = 0$ for $0 < k < n-1$. Hence, for $1 < k < n-1$,
the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
and we conclude that $H_{k}(\calT) = 0$ for $0 < k < n-1$.
Finally, the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{n}(\calT \oplus T) \longrightarrow H_{n}(\calT') \longrightarrow H_{n-1}(\calS) \longrightarrow H_{n-1}(\calT \oplus T) \longrightarrow H_{n-1}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}

% Here, if $n=1$, it already follows that $H_{n}(\calT \oplus T) \simeq H_{n}(\calT')$.
% % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq \bbZ$ because the respective simplicial complexes are path-connected. 
% Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
% one easily sees that $H_{0}(\calT') \simeq \bbZ$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
Since $\calT$ and $T$ are triangulations of manifolds with non-empty boundary, their $n$-th simplicial homology groups vanish. 
then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ are trivial and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq \bbZ$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

In the case that $\calS$ is isomorphic to a disk, we have $H_{n-1}(\calS) = 0$,
and then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T) = 0$ are trivial 
and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq {G}$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

We conclude that the simplicial homology groups of a shellable $n$-dimensional simplicial complex $\calT$ that triangulates a manifold satisfy 
\begin{align*}
    H_0(\calT) \simeq \bbZ, 
    \quad 
    H_1(\calT) \simeq \dots \simeq H_{n-1}(\calT) \simeq 0.
\end{align*}
Moreover, $H_n(\calT) \simeq {G}$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise. 
Moreover, $H_n(\calT) \simeq \bbZ$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise.

% then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follow immediately.
% Moreover, Thus $H_{n}(\calT') = H_{n}(\calT) \oplus H_{n}(\calS)$.
% We conclude that any shellable simplicial complex has the homology groups of a disk or a sphere. 
% NB: the zeroth homology is always free, and the n-th homology group, being the kernel of a mapping free abelian group, is free as well.
% \begin{align*}
%     \begin{CD}
%         0 \longrightarrow H_{1}(\calT) \oplus H_{1}(T) \longrightarrow H_{1}(\calT') \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
%     \end{CD}
% \end{align*}

\begin{remark}    
    % % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq {G}$ because the respective simplicial complexes are path-connected. 
    % Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
    % one easily sees that $H_{0}(\calT') \simeq {G}$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$.

    % then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follow immediately.
    % Moreover, Thus $H_{n}(\calT') = H_{n}(\calT) \oplus H_{n}(\calS)$.
    % We conclude that any shellable simplicial complex has the homology groups of a disk or a sphere. 
    % NB: the zeroth homology is always free, and the n-th homology group, being the kernel of a mapping free abelian group, is free as well.
    % \begin{align*}
    %     \begin{CD}
    %         0 \longrightarrow H_{1}(\calT) \oplus H_{1}(T) \longrightarrow H_{1}(\calT') \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    %     \end{CD}
    % \end{align*}
\end{remark}

>>>>

Given any simplex $T$, we write $\subsimplex(T)$ for the simplicial complex that contains all subsimplices of $T$. 
More specifically, $\subsimplex_{{k}}(T) \subseteq \subsimplex(T)$ denotes the set of $k$-dimensional subsimplices of $T$. 
Whenever $\calT$ is a simplicial complex and $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$.
We write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$. 
Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. 
In practice, we not always distinguish between points and singleton simplices. % TODO: see whether being formally correct is possible 
Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (codimension $1$ members) of this triangulation.

@@

The reason of our interest in shellable simplicial complexes is that they can be constructed by successive adhesion of simplices to intermediate complexes, 
such that the interface is well-behaved: it is a simplicial disk or a simplicial sphere. 
That leads to the following important consequence.

% TODO: we only need some basic properties, namely that the construction maintains manifolds with boundary.
% TODO: copy and adapt the content of Ziegler and Kozlov as necessary
% The notion of shelling is taken from Ziegler's book~\cite{ziegler2012lectures}

\begin{lemma}    
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, 
    then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \geq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

    Let $F$ be such a face. 
    By assumption, $F$ must be contained in exactly one $n$-simplex of $T_{0}, T_{1}, \dots, T_{m}$,
    and $F$ is in the boundary of $X_{m}$. We conclude that $I_{m}$ triangulates a submanifold of the boundary of $X_{m}$.

    On the one hand, 
    if $I_{m+1}$ is the entire boundary of $T_{m+1}$, 
    then it must also be the boundary of the topological $n$-ball $X_{m}$. 
    Hence $X_{m} \cup T_{m+1}$ must be a topological $n$-sphere and thus a manifold without boundary, which requires $m = M$.
    On the other hand, 
    if $I_{m+1}$ is a proper subset of the entire boundary of $T_{m+1}$, 
    then $X_{m} \cup T_{m+1}$ is still a topological $n$-ball.
\end{proof}

@@

\begin{lemma}    
    Any simplicial $2$-disk is shellable.
    Any simplicial $2$-sphere is shellable. 
\end{lemma}
\begin{proof}
    First, 
    let $\calS$ be any triangulation of a $2$-sphere. 
    By removing any triangle $S \in \calS$, we obtain a triangulation $\calT$ of a $2$-disk.
    Any shelling of that disk can be extended to a shelling of $\calS$ by re-inserting the first triangle $S$.
    So it remains to show that any triangulation $\calT$ of the two-dimensional disk is shellable. 
    We will construct the shelling in reverse.

    Write $M = |\calT|$. 
    We call a triangle $T \in \calT$ \emph{good} if it intersects the boundary $\partial M$ in a non-empty union of edges. 
    We show by an induction argument over the number of triangles that every triangulation of a $2$-disk contains a good triangle. 
    Clearly, this is the case if the triangulation contains only one triangle. 
    Now suppose the induction claim is true when the triangulation includes $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
    There exists a triangle $T' \in \calT$ with at least one edge in the boundary. 
    If $T'$ is good, then we are already done. 
    Otherwise, $T'$ intersects the boundary at one edge and its opposing vertex.
    Removing $T'$ splits the manifold into two pieces, each of which is a topological $2$-disk.
    By the induction assumption, they contain a good triangle.

    Since $\calT$ triangulates a $2$-disk,
    it contains a good triangle $T$. 
    If $T$ has three edges in the boundary, then $T = M$ and we are trivially done. 
    If $T$ intersects with the boundary in exactly one or two edges, 
    then $\overline{M \setminus T}$ is still a topological $2$-disk.
    Letting $\calT'$ be the triangulation obtained by removing $T$,
    we have another triangulation of the $2$-disk that intersects $T$ only at either two or one edges.
    Any shelling of $\calT'$ can be extended to a shelling of $\calT$.
%     Let $\calT$ be a simplicial $2$-disk. 
%     If $\calT$ contains only one triangle, then there is nothing to show. 
%     Otherwise, there exists at least one triangle $T$ that touches the boundary of the disk at either one or two edges 
%     {\color{red}Needs proof}.

%     Now let $\calS$ be any simplicial $2$-sphere. 
%     By removing any triangle $S \in \calS$, we obtain a simplicial $2$-disk, which is shellable. 
%     Any of its shellings can be extended to a shelling of $\calS$ by re-inserting the first triangle $S$.
\end{proof}

@@

    It remains to consider the case where $S$ is a boundary vertex. 
    Then the triangles of $\patch_{\calT}(S)$ that do not contain $V$ constitute a simplicial $2$-disk 
    which is shellable,
    and any such shelling can be extended to a shelling of $\patch_{\calT}(S)$.
\end{proof}

\begin{example}
    Not all trianguable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler2012lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions.

    For example, there exists a non-shellable triangulation of a tetrahedron. 
    If we extend this triangulation to a triangulation of a hypertetrahedron by suspending it from a new point $v_\star$, 
    then the resulting new triangulation is non-shellable and coincides with the patch around $v_\star$.
    This demonstrates that patches around boundary simplices are not necessarily shellable when the dimension exceeds $3$.
\end{remark}
